--- 
title       : World Religions Since 1945
subtitle    : ARDA World Relgion Dataset
author      : edettman (dettmanej@gmail.com)
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : default      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
--- 
<style>
.title-slide {
  background-color: #FFFFFF; /* #EDE0CF; ; #CA9F9D*/
}
em {
  font-style: italic
}

</style>

## ARDA Dataset

- World Religion Dataset (WRD) characterizes religion by country since 1945
- Data in five year increments
- Sources included census data and specific estimates (Some imputed)
- [Shinyapp](https://edettman.shinyapps.io/project) looks at the total fractions for each major religion by country and illustrates overall distributions of religions and trends in growth
- https://edettman.shinyapps.io/project

```{r, echo = F, message = F, warning = F}

library(googleVis)
require(reshape2)
require(ggplot2)
require(stringr)
setwd('c:/religion')

# Load in the information for each column in the dataset
# then clean the data to make it easier to look at
# extract the width of the columns in the ascii fixed width files to read it in
# then turn into a dataframe with variable names
info <- readLines("World Religion Dataset - National Religion Dataset.COL")
info <- info[-1]

info <- gsub('\\d{1,2}\\) {1}', '', info)

a <- str_extract(info, '^[:alpha:]+')
b <- str_extract(gsub('^[[:alpha:]3]+: ', '', info, perl = T), '\\d{1,3}')
c <- str_extract(gsub('^[[:alpha:]3]+: ', '', info, perl = T), '\\d{1,3}$')

info <- data.frame(var = a, start = as.integer(b), end = as.integer(c))

info$len <- info$end - info$start + 1

# read in the ascii fwf with the widths from the info file
# some of the columns do not match the specification, had to use fill = T 
relig <- read.fwf("World Religion Dataset - National Religion Dataset.DAT", widths = info$len, fill = T)

names(relig) <- info$var

## Turn the values that were brought in with commas to remove them
## Then can change into numeric
for(i in c(1, 7:84)){
    relig[,i] <- as.numeric(gsub(',', '', as.character(relig[,i])))
}

## Bring in the variable descriptions
desc <- readLines('World Religion Dataset - National Religion Dataset cb_descr.TXT')
desc <- data.frame(matrix(desc, ncol = 2, byrow = T))
desc <- desc[seq(2,174, by = 2), ]
colnames(desc) <- c('Var', 'Descrip')
rownames(desc) <- 1:nrow(desc)
desc[,1] <- gsub('\\d{1,2}) ', '', desc[,1])

# indices of the total numbers along with the countries
# create new dataframes for easier plotting of the total numbers of adherents
# by year
idx <- c(1:6, which(grepl('Total number', desc[,2])), 43 )
desc2 <- desc[idx,]
relig2 <- relig[,idx]

totals <- sapply(7:23, function(x) tapply(relig2[,x], relig2$YEAR, sum, na.rm = T) )

totals <- data.frame(totals)
names(totals) <- names(relig2)[7:23]

# Also calculate fractions
fractions <- totals[,1:16]/totals[,17]

totals$YEAR <- rownames(totals)

neat <- melt(data = totals, id.vars = 'YEAR')
neat$variable <- str_extract(desc[,2][match((neat$variable), desc[,1])], pattern = "^[[:alpha:]\\'-]+")

desc[c(1,6, 12, 17, 25, 29, 30, 31),2]

```

--- .class #id 

## Growth of Religious Populations Since 1945  

```{r, echo = F, fig.height = 5.5, fig.width = 12}

qplot(x = YEAR, y = value, group = factor(variable), 
      data = neat[neat$variable != 'Total',], colour = factor(variable), geom = 'line', size = I(1.5)) +
      annotate('text', x = factor(c(1975, 2002, 2002)),
               y = c(1.5e9, 1.5e9, 1.0e9), 
               label = c('Christianity', 'Islam', 'Hindu')) +
      ylab('Population') + ggtitle('Growth of Populations of Religions')  

```

- Above line graph shows the overall populations of each indicated religion since 1945 
- ARDA dataset shows these populations at the level of the individual country 
- This [Shinyapp](https://edettman.shinyapps.io/project) can be used to investigate these trends at a more granular level

---
## World Geographic Plot of Individual Religions  
#### Christianity in 2010
```{r, echo = F, results = 'asis', message = F}

# now just isolate the fractions for each country by year
# indices of the total fractions along with the countries
idx <- c(1:6, which(grepl('Total percentage', desc[,2])) )

desc <- desc[idx,]
relig <- relig[,idx]

# choices for religion
a <- str_extract(desc[7:nrow(desc),2], '^[^:]+')

# choices for year are 1945 to 2010 in 5 year increments
# choice1 and choice2 are used as user inputs in the shiny app
choice1 = "Christianity"
choice2 = 2010
idx <- which(a == choice1) + 6
partRelig <- relig[relig$YEAR == choice2, c(6,idx)]
names(partRelig) <- c('Country', 'Freq')
G <- gvisGeoChart(partRelig,"Country", "Freq",
                  options = list(title = 'Christianity in 2010', 
                                 colorAxis = "{colors: ['#00853f', 'black', '#e31b23']}",
                                 datalessRegionColor = '#ffffff',
                                 height = 300, width = 600))
print(G, 'chart')

```
- On the [Shinyapp](https://edettman.shinyapps.io/project), the user can select an individual religion and year
- A heatmap of the fraction of people adhering to that relgion by year is then displayed in the *Map* tab
- Green is low, black is middle, red is high, and white is missing data
- Hover over individual countries to see the fraction of that religion

---
## Trends in Individual Countries

```{r, echo = F, fig.height = 4, fig.width = 8, fig.align='center'}
# Build a graphing function to show this for specific countries
temp <- melt(relig[,c(1,6:22)], id.vars = c('YEAR', 'COUNTRY'))
temp$variable <- str_extract(desc[,2][match(temp$variable, desc[,1])], pattern = "^[[:alpha:]\\'-]+") 
temp$COUNTRY <- gsub('^[ ]+', '', temp$COUNTRY)


pick = 'China'
choice2 = 'Syncretic'

qplot(x = YEAR, y = value,  fill = I('firebrick1'),
      data = temp[temp$COUNTRY == pick & temp$variable == choice2,], 
      geom = 'bar', stat= 'identity') +
      ylab('Fraction') + ggtitle(paste('Growth of', choice2, 'Religion in', pick)) 

```

- On the [Shinyapp](https://edettman.shinyapps.io/project), the user can also select an individual country
- _Growth_ tab will show the trends in the selected relgion in that country between 1945 and 2010
- A third _Table_ tab shows a sortable and searchable table of all countries in the selected year for the selected religion